---
title: "Francisco_Group_Proj"
author: "Taylor Francisco"
date: "2023-11-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
rm(list=ls())
```


#Set Up
##Load Packages
```{r}
library(tidyverse)
library(dplyr)
library(purrr)
library(readstata13)
library(plyr)
library(stargazer)
library(stringr)
library(ggplot2)
library(psych)
library(plm)
library(visreg)
library(VGAM)
library(car)
```


##Upload Data
```{r}
anes_df <- read_csv("anes_timeseries_cdf_csv_20220916.csv")
vote_df <- read_csv("Turnout_1980_2022_v1.0.csv")
abortion_policy_df <- read_csv("Abortion_Policies.csv")
```

#Clean Data

##ANES dataset
```{r}
#select only variables of interest

anes_select_df <- anes_df %>% select("VCF0302", "VCF0705", "VCF0748", "VCF0900", 
                                    "VCF0901b", "VCF9234", "VCF9235", "VCF0004", 
                                    "VCF0006", "VCF0101", "VCF0102", "VCF0104", 
                                    "VCF0105a", "VCF0105b", "VCF0106","VCF0107",
                                    "VCF0108", "VCF0110","VCF0112", "VCF0128", 
                                    "VCF0132","VCF0133", "VCF0138", "VCF0139", 
                                    "VCF0140a","VCF0142", "VCF0900b", "VCF0230", 
                                   "VCF0837", "VCF0838")

#save as a csv for faster upload later
#write.csv(anes_select_df, file ="ANES_select.csv", row.names = FALSE)

#rename variables with descriptive name
##note name STATE, STATE_ABV, YEAR all caps to match other dataframes

                                            
 d <- anes_select_df     
 d$vote_pres = d$VCF0705
 d$vote_elect_day =  d$VCF0748
 d$congress_dist  = d$VCF0900
d$STATE_ABV = d$VCF0901b 
d$dem_cand_abort = d$VCF9234 
 d$rep_cand_abort = d$VCF9234
 d$YEAR = d$VCF0004
 d$resp_id = d$VCF0006
d$age = d$VCF0101
d$age_group = d$VCF0102
d$gender  = d$VCF0104
d$race_seven  = d$VCF0105a 
d$race_four   = d$VCF0105b 
d$race_three  = d$VCF0106 
d$hisp  = d$VCF0108
d$edu_four  = d$VCF0110
d$edu_seven   = d$VCF0140a 
d$region  = d$VCF0112
d$religion  = d$VCF0128 
d$num_child = d$VCF0139 
d$birth_place = d$VCF0142 
d$state_congress_dist = d$VCF0900b
d$anti_abort_therm  = d$VCF0230
d$when_abort  = d$VCF0837
d$abort_law   = d$VCF0838


#anes_select_df <- anes_select_df %>% rename(party_id = VCF0302,
#                           vote_pres = VCF0705,
#                     vote_elect_day =  VCF0748, 
#                      congress_dist  = VCF0900, 
#                        STATE_ABV = VCF0901b, 
#                      dem_cand_abort = VCF9234, 
#                      rep_cand_abort = VCF9234, 
#                                YEAR = VCF0004, 
#                            resp_id = VCF0006, 
#                                age = VCF0101, 
#                          age_group = VCF0102,
#                            gender  = VCF0104, 
#                       race_seven  = VCF0105a, 
#                       race_four   = VCF0105b, 
#                       race_three  = VCF0106, 
#                             hisp  = VCF0108, 
#                         edu_four  = VCF0110,
#                       edu_seven   = VCF0140a, 
#                           region  = VCF0112, 
#                         religion  = VCF0128, 
#                         num_child = VCF0139, 
#                       birth_place = VCF0142, 
#               state_congress_dist = VCF0900b,
#                 anti_abort_therm  = VCF0230,
#                       when_abort  = VCF0837,
#                       abort_law   = VCF0838)

anes_renamed_df <- d %>% select("vote_pres", "vote_elect_day", "congress_dist",  
              "STATE_ABV", "dem_cand_abort", "rep_cand_abort", "YEAR", "resp_id", 
              "age", "age_group", "gender", "race_seven", "race_four", "race_three",
              "hisp", "edu_four", "edu_seven", "region", "religion", "num_child",
              "birth_place", "state_congress_dist", "anti_abort_therm", "when_abort",
              "abort_law") 
  
 

#save as a csv for faster upload later
#write.csv(anes_renamed_df, file ="ANES_select.csv", row.names = FALSE)



```

##Voting dataset

```{r}
#get rid of US as a state- doesn't contain any information just a header interspersed with the data
vote_df <- vote_df %>% filter(STATE != "United States")
unique(vote_df$STATE_ABV) #no more cleaning here

```
##Abortion Policy dataset
```{r}
#change State to STATE to allow for merging by state later on
abortion_policy_df <- abortion_policy_df %>% mutate(STATE = State)
#need to add STATE_ABV to merge with other two datasets
state <- read_csv("states.csv")
state <- state %>% mutate(STATE = State)

abortion_policy_df <- left_join(abortion_policy_df, state, by = "STATE")

#rename abbreviation to match other datasets 
abortion_policy_df <- abortion_policy_df %>% mutate(STATE_ABV = Abbreviation)
#since the mutate function keeps old state columns, get rid of for simplicity
abortion_policy_df <- abortion_policy_df %>% select("Status of Abortion",
                                                    "Changed","State Law",
                                                    "STATE", "STATE_ABV")

```

##Combine dataframes for analysis
```{r}
#start by combining anes and voter datasets
anes_vote_df <- left_join(anes_renamed_df, vote_df, by = c("STATE_ABV", "YEAR"))
#expect it to be as long as anes as there is more individual obvs on that as vote is just info by state and year
#add in abortion policy df

combined_df <- left_join(anes_vote_df, abortion_policy_df, by = c("STATE_ABV"))
#expecting just a few more columns to be added to above 

write.csv(combined_df, file ="combined_data.csv", row.names = FALSE)
```


#Analysis
get summary stats by state and year, save as new df
use new df above in regression with unit-level being state
this will only be up to 2020
DID would be trying to get a bit of inight to see if the trend would be effected by Dobbs
##Exploratory regression 
##Differences-in-differences 

```{r}
#create a variable with the year of "treatment" which is 2022
df$time = ifelse(df$YEAR >= 2022, 1, 0)

#create interaction between time and the treated variable= changed 
df$did = df$time * df$Changed

#Estimating the DID estimator
didreg = lm(VEP ~ Changed + time + did, data = df)
summary(didreg)

```
 The coefficient for ‘did’ is the differences-in-differences
estimator, not significant in this case, but a positive direction. 

```{r}
didreg1 = lm(VEP ~ Changed*time, data = df)
summary(didreg1)
```
The coefficient for ‘treated#time’ is the differences-indifferences estimator 
(did in last estimate). Still not significant. 

```{r}
#simple model 
reg = lm(VEP ~ Changed, data = df)
summary(reg)
```
still not significant

#some exploratory graphs

```{r}
#code from https://lfoswald.github.io/2021-spring-stats2/materials/session-8/08-online-tutorial/
df %>%
mutate(Changed = ifelse(Changed == 1, "Treated (D=1)", "Untreated (D=0)")) %>%
group_by(time, Changed) %>% # group to extract means of each group at each time
mutate(group_mean = mean(VEP)) %>% # extract means of each group at each time
ggplot(., aes(x = VEP, fill = factor(Changed))) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(name = " ", # changes to fill dimension
                     values = c("#cc0055", "#a7a8aa"),
                     labels = c("Treatment", "Control")) +
  facet_grid(Changed~time) + # we specify the matrix (treatment and period)
  geom_vline(aes(xintercept = group_mean), linetype = "longdash") + # add vertical line with the mean
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "VEP",
       y = "Density")
```
```{r}
#code from https://lfoswald.github.io/2021-spring-stats2/materials/session-8/08-online-tutorial/
df %>%
group_by(time, Changed) %>% # group to extract means of each group at each time
mutate(group_mean = mean(VEP)) %>%
ggplot(aes(x = time, y = group_mean, color = factor(Changed))) +
  geom_point() +
  geom_line(aes(x = time, y = group_mean)) +
  scale_x_continuous(breaks = c(0,1)) +
  scale_color_manual(name = " ", # changes to color dimension
                     values = c("#a7a8aa", "#cc0055"),
                     labels = c("Control", "Treatment")) +
  labs(x = "Time periods", y = "VEP", color = "Treatment group")+
  theme_minimal() 
```
Would see seperate lines with different slopes if they were different. 

