---
title: "Francisco_Group_Proj"
author: "Taylor Francisco"
date: "2023-11-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
rm(list=ls())
```


#Set Up
##Load Packages
```{r}
library(tidyverse)
library(dplyr)
library(purrr)
library(readstata13)
library(plyr)
library(stargazer)
library(stringr)
library(ggplot2)
library(psych)
library(plm)
library(visreg)
library(VGAM)
library(car)
```


##Upload Data
```{r}
anes_df <- read_csv("anes_timeseries_cdf_csv_20220916.csv")
vote_df <- read_csv("Turnout_1980_2022_v1.0.csv")
abortion_policy_df <- read_csv("Abortion_Policies.csv")
```

#Clean Data

##ANES dataset
```{r}
#select only variables of interest

anes_select_df <- anes_df %>% select("VCF0302", "VCF0705", "VCF0748", "VCF0900", 
                                    "VCF0901b", "VCF9234", "VCF9235", "VCF0004", 
                                    "VCF0006", "VCF0101", "VCF0102", "VCF0104", 
                                    "VCF0105a", "VCF0105b", "VCF0106","VCF0107",
                                    "VCF0108", "VCF0110","VCF0112", "VCF0128", 
                                    "VCF0132","VCF0133", "VCF0138", "VCF0139", 
                                    "VCF0140a","VCF0142", "VCF0900b", "VCF0230", 
                                   "VCF0837", "VCF0838")

#save as a csv for faster upload later
#write.csv(anes_select_df, file ="ANES_select.csv", row.names = FALSE)

#rename variables with descriptive name
##note name STATE, STATE_ABV, YEAR all caps to match other dataframes
 d <- anes_select_df  #change name to d to simplify renaming since tidyverse rename() wasn't working
 d$vote_pres = d$VCF0705
 d$vote_elect_day =  d$VCF0748
 d$congress_dist  = d$VCF0900
d$STATE_ABV = d$VCF0901b 
d$dem_cand_abort = d$VCF9234 
 d$rep_cand_abort = d$VCF9234
 d$YEAR = d$VCF0004
 d$resp_id = d$VCF0006
d$age = d$VCF0101
d$age_group = d$VCF0102
d$gender  = d$VCF0104
d$race_seven  = d$VCF0105a 
d$race_four   = d$VCF0105b 
d$race_three  = d$VCF0106 
d$hisp  = d$VCF0108
d$edu_four  = d$VCF0110
d$edu_seven   = d$VCF0140a 
d$region  = d$VCF0112
d$religion  = d$VCF0128 
d$num_child = d$VCF0139 
d$birth_place = d$VCF0142 
d$state_congress_dist = d$VCF0900b
d$anti_abort_therm  = d$VCF0230
d$when_abort  = d$VCF0837
d$abort_law   = d$VCF0838

#save a df with only renamed variables, dropping VCF#### names
anes_renamed_df <- d %>% select("vote_pres", "vote_elect_day", "congress_dist",  
              "STATE_ABV", "dem_cand_abort", "rep_cand_abort", "YEAR", "resp_id", 
              "age", "age_group", "gender", "race_seven", "race_four", "race_three",
              "hisp", "edu_four", "edu_seven", "region", "religion", "num_child",
              "birth_place", "state_congress_dist", "anti_abort_therm", "when_abort",
              "abort_law") 
  
 

#save as a csv for faster upload later
#write.csv(anes_renamed_df, file ="ANES_select.csv", row.names = FALSE)

#check that there are only 50 states
unique(anes_renamed_df$STATE_ABV) #Note, DC and 99 need to be removed
#get rid of DC and 99 
anes_renamed_df <- anes_renamed_df %>% filter(STATE_ABV != "DC")
anes_renamed_df <- anes_renamed_df %>% filter(STATE_ABV != "99")
unique(anes_renamed_df$STATE_ABV)#no more cleaning here, only 50 states

unique(anes_renamed_df$YEAR)
#vote_df only has years 1980 and above (our years of interest), so remove earlier years

anes_renamed_df <-  anes_renamed_df %>% filter(YEAR > 1980)
unique(anes_renamed_df$YEAR) #only 1982 onwards to 2020 
```

##Voting dataset

```{r}
#check that there are only 50 states
unique(vote_df$STATE) #note District of Columbia included. also Pennsylvania* has a * added?

#get rid of US as a state- doesn't contain any information just a header interspersed with the data
vote_df <- vote_df %>% filter(STATE != "United States")
vote_df <- vote_df %>% filter(STATE_ABV != "DC")
unique(vote_df$STATE_ABV) #no more cleaning here, only 50 states 

#change VEP_TURNOUT_RATE to be numberical, right now has % which makes it character
vote_df$VEP_TURNOUT_RATE <- as.numeric(sub("%", "", vote_df$VEP_TURNOUT_RATE))
class(vote_df$VEP_TURNOUT_RATE) #is now numeric
```

##Abortion Policy dataset
```{r}
#change State to STATE to allow for merging by state later on
abortion_policy_df <- abortion_policy_df %>% mutate(STATE = State)
#need to add STATE_ABV to merge with other two datasets
state <- read_csv("states.csv") #dataframe with just State name and State abbreviation
#change State to STATE to allow for merging by state later on
state <- state %>% mutate(STATE = State)

abortion_policy_df <- left_join(abortion_policy_df, state, by = "STATE")

#rename abbreviation to match other datasets 
abortion_policy_df <- abortion_policy_df %>% mutate(STATE_ABV = Abbreviation)
#since the mutate function keeps old state columns, get rid of for simplicity
abortion_policy_df <- abortion_policy_df %>% select("Status of Abortion",
                                                    "Changed","State Law",
                                                    "STATE", "STATE_ABV")
#rename variables to remove spaces for ease of analysis later on

abortion_policy_df$abort_status = abortion_policy_df$`Status of Abortion`
abortion_policy_df$state_law  = abortion_policy_df$`State Law`
```

##Combine dataframes for analysis
```{r}
#start by combining anes and voter datasets
anes_vote_df <- left_join(anes_renamed_df, vote_df, by = c("STATE_ABV", "YEAR"))
#expect it to be as long as anes as there is more individual obvs on that as vote is just info by state and year


#add in abortion policy df
combined_df <- left_join(anes_vote_df, abortion_policy_df, by = c("STATE_ABV"))
#expecting just a few more columns to be added to above 

#need just a dataframe of voting and state policy for DiD analysis (anes does not contain 2022)
vote_abort_df <- left_join(vote_df, abortion_policy_df, by = c("STATE_ABV"))

#save as csv to send to other group members cleaned data 
#write.csv(combined_df, file ="combined_data.csv", row.names = FALSE)
#write.csv(vote_abort_df, file ="vote_abort.csv", row.names = FALSE)
```

#Descriptive Stats
##Determine missingness of dataframe
```{r}
combined_df %>% is.na() %>% colSums()
#for what we are interested in 
final_var_df <- combined_df %>% select("STATE_ABV", "YEAR", "age", "gender", 
                                        "religion", "anti_abort_therm", 
                                        "VEP_TURNOUT_RATE", "Changed",
                                        "abort_status", "state_law", "race_four", "abort_law")

final_var_df %>% is.na() %>% colSums()
#only ones missing are VEP_TURNOUT_RATE and ANTI_ABORT_THERM
#VEP_TURNOUT_RATE 
  7775/42348 # only 0.1835978 missing, not a big deal
#anti_abort_therm
  36071/42348 #a lot more of this is missing
#abort_law
  1510/42348

#find out how much data is missing when using abort_law
detach(package:plyr)
 missing_values_count <- final_var_df %>%
  group_by(STATE_ABV, YEAR) %>%
  summarize(
    mean_value = mean(abort_law, na.rm = TRUE),
    missing_count = sum(is.na(abort_law))
  )
 
  missing_values_count

#abort_law seems like the way to go
```
##Descriptive stats for 
```{r}
#basic stats for all variables in the whole dataset
sub <- combined_df[,c("age", "gender", "religion","anti_abort_therm", "VEP_TURNOUT_RATE", "Changed",
                                        "abort_status", "state_law", "race_four", "abort_law")]


sum <- describe(sub)
sum
```


```{r}
all_stats <- final_var_df %>% split(final_var_df$STATE_ABV, final_var_df$YEAR) %>% map(summary)
states_stats  <- final_var_df  %>% split(final_var_df$STATE_ABV) %>% map(summary)
```

##Stats by State and Year 
```{r}
#what i need to change to be state level 
#"age", "gender", "religion","race_four", "abort_law"

#age continuous -- mean 
age_stats <- combined_df %>%                            
 group_by(STATE_ABV, YEAR) %>%
 summarize(mean_age = mean(age, na.rm = TRUE),
           sd_age =sd(age)
          )

#gender binary --- mean 
gender_stats <- combined_df %>%                            
 group_by(STATE_ABV, YEAR) %>%
 summarize(mean_gender = mean(gender, na.rm = TRUE),
           sd_gender =sd(gender)
          )
#religion categorical--- mode
##First create mode function 
my_mode <- function(codes){
  which.max(tabulate(codes))
}
##calculate stats
religion_stats <- combined_df %>%
  group_by(STATE_ABV, YEAR) %>%
  summarise(mode_religion = my_mode(religion))
unique(religion_stats$mode_religion)
#race factor =----------- mode
race_stats <- combined_df %>%
  group_by(STATE_ABV, YEAR) %>%
  summarise(mode_race = my_mode(race_four))
unique(race_stats$mode_race)
#abort_law is a factor----mode
abort_stats <- combined_df %>%
  group_by(STATE_ABV, YEAR) %>%
  summarise(mode_abort = my_mode(abort_law))
unique(abort_stats$mode_abort)
abort_stats

```
#Analysis
##changing data to be state-level
```{r}
#merge the stats to be included in main df to be used for state-level unit analysis
state_level_df <- left_join(combined_df, age_stats, by = c("STATE_ABV", "YEAR"))
state_level_df <- left_join(state_level_df, gender_stats, by = c("STATE_ABV", "YEAR"))
state_level_df <- left_join(state_level_df, religion_stats, by = c("STATE_ABV", "YEAR"))
state_level_df <- left_join(state_level_df, race_stats, by = c("STATE_ABV", "YEAR"))
state_level_df <- left_join(state_level_df, abort_stats, by = c("STATE_ABV", "YEAR"))
```

##Simple regression
```{r}
unique(state_level_df$mode_abort)
state_level_df$abort.fact <- factor(state_level_df$mode_abort, levels = c(1, 2, 3, 4), 
                        labels = c("Never", "Only for rape, incest, or danger to Mother",
                                   "Allowed for other reasons", 
                                   "Always"), exclude = NA, ordered= FALSE)

#only viewing thoughts on abortion
m1 <- lm(VEP_TURNOUT_RATE ~ abort.fact, data =state_level_df)
summary(m1)

#adding in some controls 
m2 <- lm(VEP_TURNOUT_RATE ~ abort.fact + YEAR + STATE_ABV, data =state_level_df)
summary(m2)

#add in interaction with State 

m3 <- lm(VEP_TURNOUT_RATE ~ abort.fact + YEAR + STATE_ABV + abort.fact*STATE_ABV, data =state_level_df)
summary(m3)

```

##Regression with demo as controls
```{r}
m4 <- lm(VEP_TURNOUT_RATE ~ abort.fact + YEAR + mean_age + mean_gender + 
           mode_religion + mode_race, data =state_level_df)
summary(m4)
```

##Differences-in-differences 

```{r}
#create a variable with the year of "treatment" which is 2022
vote_abort_df$time = ifelse(vote_abort_df$YEAR >= 2021, 1, 0)
unique(vote_abort_df$YEAR)
unique(vote_abort_df$time)

#create interaction between time and the treated variable= changed 
vote_abort_df$did = vote_abort_df$time * vote_abort_df$Changed

#Estimating the DID estimator
didreg = lm(VEP_TURNOUT_RATE ~ Changed + time + did, data = vote_abort_df)
summary(didreg)

```
 

```{r}
didreg1 = lm(VEP_TURNOUT_RATE ~ Changed*time, data = vote_abort_df)
summary(didreg1)
```
The coefficient for ‘treated#time’ is the differences-indifferences estimator 
(did in last estimate). Still not significant. 

```{r}
#simple model 
reg = lm(VEP_TURNOUT_RATE ~ Changed, data = vote_abort_df)
summary(reg)
```
still not significant

#some exploratory graphs

```{r}
#code from https://lfoswald.github.io/2021-spring-stats2/materials/session-8/08-online-tutorial/
vote_abort_df %>%
mutate(Changed = ifelse(Changed == 1, "Treated (D=1)", "Untreated (D=0)")) %>%
group_by(time, Changed) %>% # group to extract means of each group at each time
mutate(group_mean = mean(VEP_TURNOUT_RATE)) %>% # extract means of each group at each time
ggplot(., aes(x = VEP_TURNOUT_RATE, fill = factor(Changed))) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(name = " ", # changes to fill dimension
                     values = c("#cc0055", "#a7a8aa"),
                     labels = c("Treatment", "Control")) +
  facet_grid(Changed~time) + # we specify the matrix (treatment and period)
  geom_vline(aes(xintercept = group_mean), linetype = "longdash") + # add vertical line with the mean
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "VEP Turnout Rate",
       y = "Density")
```
```{r}
#code from https://lfoswald.github.io/2021-spring-stats2/materials/session-8/08-online-tutorial/
vote_abort_df%>%
group_by(YEAR, Changed) %>% # group to extract means of each group at each time
mutate(group_mean = mean(VEP_TURNOUT_RATE)) %>%
ggplot(aes(x = YEAR, y = group_mean, color = factor(Changed))) +
  geom_point() +
  geom_line(aes(x = YEAR, y = group_mean)) +
  scale_color_manual(name = " ", # changes to color dimension
                     values = c("#a7a8aa", "#cc0055"),
                     labels = c("Control", "Treatment")) +
  labs(x = "Time periods", y = "VEP Turnout Rate", color = "Treatment group")+
  theme_minimal() 
```
Would see seperate lines with different slopes if they were different. 
#another graph 
```{r}

vote_abort_df %>%
  group_by(Changed, time) %>%
  summarize(mean_outcome = mean(VEP_TURNOUT_RATE),
            ci_lower = mean_outcome - 1.96 * sd(VEP_TURNOUT_RATE) / sqrt(n()),
            ci_upper = mean_outcome + 1.96 * sd(VEP_TURNOUT_RATE) / sqrt(n())) %>%
  
  ggplot(aes(x = time, y = mean_outcome, color = as.factor(Changed))) +
  geom_line(aes(group = Changed), size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.1) +
  
  labs(title = "Difference-in-Differences Plot",
       x = "Time",
       y = "Mean Outcome") +
  
  theme_minimal()
```


